---
title: "Data Science Homework 3"
author: Kristina Howell
output: github_document
---

The following settings and libraries will be used throughout the homework. 

```{r settings, message = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "virirdis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Load the datasets from the p8105 library. 

```{r}
library(p8105.datasets)
data("instacart")
```

### Dataset Description

The dataset **instacart** contains information about specific orders from the instacart online grocery service with 1,384,617 observations from 131,209 unique users. Key variables in the dataset include _department_, _aisle_, _product_name_, and _order_id_, as well as various entries representing time, day of week, and days since prior order. 

Overall, the dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Each column represents an individual item ordered from a user. For example, the first 8 rows represent:

* the _order_id_ 1 
* from _user_id_ 112108 
* includes items from _department_ "dairy eggs", "produce", and "canned goods"
* featuring _product_name_ "Bulgarian Yogurt" ( _product_id_: 49302 ) and _product_name_ "Organic Celery Hearts" ( _product_id_: 10246 )

The **complete list of variables** in the dataset is as follows: `r names(instacart)`.

### Problem 1 Questions


















